---
title: "Reading Cell Count data from spreasheet (XL)"
output: html_notebook
---
PURPOSE: This script reads in cell count data from excel spreadsheet

1. What are the sampling dates?
2. What are the sample IDs for each year/site?
3. What is the sampling depth for each sampling date?
4. What are the t-1 measurements for each sample? 

Where are we missing data?

File History: 
[23MAR21] this file was copied from the Bandelier Project: CellCountDatabase_00.Rmd
```{r}
library(readxl)
library(TDPanalysis)
library(lubridate)
library(stringr)

# clear workspace
rm(list=ls())

#fname_current <- "CellCounts_BAN_10JUN19_Rtest.xlsx"
#pathname1 <- file.path("~/Documents/Xylogenesis/Project_5 Bandelier/3. R/src/data",fname_current)

fname_current <- "BIG_from 01JULY2020_2.xlsx"
pathname1 <- file.path("~/Documents/Xylogenesis/Project_6 Monson 2018/2. Data/1. Cell Counts/BIG",fname_current)

# [5JAN21] need to insert pathname
data <- read_xlsx(pathname1, 
                  sheet = "data",
                  range = "A1:AF471", 
                  na = "-999")
data <- data.frame(data)
```

colnames(data) <- c("actions", 
                    "junk", 
                    "smp_ID",
                    "junk",
                    "month_s",
                    "day_s",
                    "year_s",
                    "C1","E1","T1","M1",
                    "C2","E2","T2","M2",
                    "C3","E3","T3","M3"
                    )
```{r}
# "actions" are labelled: "RE-DO","REJECT" or "NA". For "NA" records, there is cell count data
# Parse out samples that need to be Redone or Rejected
data.RR <- subset(data,!is.na(data$actions))

# ~~~~~~~~~~~~~~~~~~  ONLY RECORDS WITH CELL COUNT DATA

# Parse out only samples with cell count data
dataCC.data <- subset(data,is.na(data$actions))
dataCC.data2 <- dataCC.data[,c(-1,-2,-4)]
dataCC.data2 <- data.frame(dataCC.data2)
nrec <- length(dataCC.data2[,1])

#data.plus1$actions[is.na(data.plus1$actions)] <- 43
#data.data <- subset(data.plus1,data.plus1$actions==43)

# CONVERT DAYS TO JULIAN DAYS
ndays_month <- c(31,28,31,30,31,30,31,31,30,31,30,31)
ndays_moncum <- cumsum(ndays_month)

doy <- ndays_moncum[dataCC.data2$month_s-1]+as.numeric(dataCC.data2$day_s)
DOY <- matrix(data=doy,nrow=nrec,ncol=1)

# add DOY to raw data set
chk1.data <- cbind(DOY,dataCC.data2)
```


PLOTTING SAMPLING DEPTH BY SAMPLING DAY
This gives me an idea of how much data I have and when.

 This chunk is for PIPO
```{r}
# PIPO
# Subset all PIPO samples - ones that begin with "B"
PIPO_01.data <- chk1.data[str_detect(chk1.data$smp_ID,"^B\\w*"),]  # string specifies ONLY IDs that begin with B 

# Sort data by sampling date: 1) yr, 2) month, 3) day
# then by 4) smp_ID
PIPOByDate.data <- PIPO_01.data[order(PIPO_01.data$year_s,
                              PIPO_01.data$month_s,
                              PIPO_01.data$day_s,
                              PIPO_01.data$smp_ID),]

# What years do we have? 
chk.SmpYrs <- unique(PIPOByDate.data$year_s)
chk.SmpYrs

#Do all dates make sense? - make histogram for each year
PPdata.2016 <- subset(PIPOByDate.data,PIPOByDate.data$year_s==16)
PPdata.2017 <- subset(PIPOByDate.data,PIPOByDate.data$year_s==17)
PPdata.2018 <- subset(PIPOByDate.data,PIPOByDate.data$year_s==18)

# ~~~~~~~~~~~~~~~ PIPO 2016
SmpDay2016_beg <- min(PPdata.2016$DOY,na.rm = T)
SmpDay2016_end <- max(PPdata.2016$DOY,na.rm = T)
SmpPer2016_days <- SmpDay2016_end-SmpDay2016_beg+1

# Sampling dates
SmpDatesPP.2016 <- unique(PPdata.2016$DOY)

# Tree IDs
TreeIDsPP.2016 <- unique(PPdata.2016$smp_ID)

hist(PPdata.2016$DOY,
     breaks=SmpPer2016_days, 
     main = "",
     xaxt="n")
abline(v=ndays_moncum, lty=2)

axis(1,at=seq(0,365,30))
axis(3, at = c(15,45,75,106,136,167,197,228,259,289,319,349),
     labels=c("J","F","M","A","M","J","J","A","S","O","N","D"),
     tck=-.00,padj=1,cex.axis=1.2)

# ~~~~~~~~~~~~~~~ PIPO 2017
SmpDay2017_beg <- min(PPdata.2017$DOY,na.rm = T)
SmpDay2017_end <- max(PPdata.2017$DOY,na.rm = T)
SmpPer2017_days <- SmpDay2017_end-SmpDay2017_beg+1

# Sampling dates
SmpDatesPP.2017 <- unique(PPdata.2017$DOY)

# Tree IDs
TreeIDsPP.2017 <- unique(PPdata.2017$smp_ID)

hist(PPdata.2017$DOY,
     breaks=SmpPer2017_days, 
     main = "",
     xaxt="n")
abline(v=ndays_moncum, lty=2)

axis(1,at=seq(0,365,30))
axis(3, at = c(15,45,75,106,136,167,197,228,259,289,319,349),
     labels=c("J","F","M","A","M","J","J","A","S","O","N","D"),
     tck=-.00,padj=1,cex.axis=1.2)

# ~~~~~~~~~~~~~~~ PIPO 2018
SmpDay2018_beg <- min(PPdata.2018$DOY,na.rm = T)
SmpDay2018_end <- max(PPdata.2018$DOY,na.rm = T)
SmpPer2018_days <- SmpDay2018_end-SmpDay2018_beg+1

# Sampling dates
SmpDatesPP.2018 <- unique(PPdata.2018$DOY)

# Tree IDs
TreeIDsPP.2018 <- unique(PPdata.2018$smp_ID)

hist(PPdata.2018$DOY,
     breaks=SmpPer2018_days, 
     main = "",
     xaxt="n")
abline(v=ndays_moncum, lty=2)

axis(1,at=seq(0,365,30))
axis(3, at = c(15,45,75,106,136,167,197,228,259,289,319,349),
     labels=c("J","F","M","A","M","J","J","A","S","O","N","D"),
     tck=-.00,padj=1,cex.axis=1.2)
```

This chunk is for PSME
```{r}
# 
# Subset all PSME samples - ones that begin with "B"
PSME_01.data <- chk1.data[str_detect(chk1.data$smp_ID,"^D\\w*"),]

# Sort data by sampling date: 1) yr, 2) month, 3) day
# then by 4) smp_ID
PSMEByDate.data <- PSME_01.data[order(PSME_01.data$year_s,
                              PSME_01.data$month_s,
                              PSME_01.data$day_s,
                              PSME_01.data$smp_ID),]

# What years do we have? 
chk.SmpYrs <- unique(PSMEByDate.data$year_s)
chk.SmpYrs

#Do all dates make sense? - make histogram for each year
DFdata.2016 <- subset(PSMEByDate.data,PSMEByDate.data$year_s==16)
DFdata.2017 <- subset(PSMEByDate.data,PSMEByDate.data$year_s==17)
DFdata.2018 <- subset(PSMEByDate.data,PSMEByDate.data$year_s==18)

# ~~~~~~~~~~~~~~~ PSME 2016
SmpDay2016_beg <- min(DFdata.2016$DOY,na.rm = T)
SmpDay2016_end <- max(DFdata.2016$DOY,na.rm = T)
SmpPer2016_days <- SmpDay2016_end-SmpDay2016_beg+1

# Sampling dates
SmpDatesDF.2016 <- unique(DFdata.2016$DOY)

# Tree IDs
TreeIDsDF.2016 <- unique(DFdata.2016$smp_ID)

hist(DFdata.2016$DOY,
     breaks=SmpPer2016_days, 
     main = "",
     xaxt="n")
abline(v=ndays_moncum, lty=2)

axis(1,at=seq(0,365,30))
axis(3, at = c(15,45,75,106,136,167,197,228,259,289,319,349),
     labels=c("J","F","M","A","M","J","J","A","S","O","N","D"),
     tck=-.00,padj=1,cex.axis=1.2)

# ~~~~~~~~~~~~~~~ PSME 2017
SmpDay2017_beg <- min(DFdata.2017$DOY,na.rm = T)
SmpDay2017_end <- max(DFdata.2017$DOY,na.rm = T)
SmpPer2017_days <- SmpDay2017_end-SmpDay2017_beg+1

# Sampling dates
SmpDatesDF.2017 <- unique(DFdata.2017$DOY)

# Tree IDs
TreeIDsDF.2017 <- unique(DFdata.2017$smp_ID)

hist(DFdata.2017$DOY,
     breaks=SmpPer2017_days, 
     main = "",
     xaxt="n")
abline(v=ndays_moncum, lty=2)

axis(1,at=seq(0,365,30))
axis(3, at = c(15,45,75,106,136,167,197,228,259,289,319,349),
     labels=c("J","F","M","A","M","J","J","A","S","O","N","D"),
     tck=-.00,padj=1,cex.axis=1.2)

# ~~~~~~~~~~~~~~~ PSME 2018
SmpDay2018_beg <- min(DFdata.2018$DOY,na.rm = T)
SmpDay2018_end <- max(DFdata.2018$DOY,na.rm = T)
SmpPer2018_days <- SmpDay2018_end-SmpDay2018_beg+1

# Sampling dates
SmpDatesDF.2018 <- unique(DFdata.2018$DOY)

# Tree IDs
TreeIDsDF.2018 <- unique(DFdata.2018$smp_ID)

hist(DFdata.2018$DOY,
     breaks=SmpPer2018_days, 
     main = "",
     xaxt="n")
abline(v=ndays_moncum, lty=2)

axis(1,at=seq(0,365,30))
axis(3, at = c(15,45,75,106,136,167,197,228,259,289,319,349),
     labels=c("J","F","M","A","M","J","J","A","S","O","N","D"),
     tck=-.00,padj=1,cex.axis=1.2)

```


NEXT STEPS
1) Adjust sampling dates
2) Compile and re-plot data by sub-group, eg. species


COMPUTE AVERAGE COUNTS
this applies to any datasort 
```{r}
# get number of records
n_rec.data <- length(data.2017)

set.data <- SortByDate.data

AvgCounts.data <- matrix(data = NA,nrow = n_rec,ncol = 4)
JD <- matrix(data = NA,nrow = n_rec,ncol = 1)
for (i in 1:n_rec){
  JD[i,1] <- floor(juldate(c((set.data$year_s[i]+2000),set.data$month_s[i],set.data$day_s[i]))) - 59214
  C <- c(set.data$C1[i],set.data$C2[i],set.data$C3[i])
  E <- c(set.data$E1[i],set.data$E2[i],set.data$E3[i])
  Th <- c(set.data$T1[i],set.data$T2[i],set.data$T3[i])
  M <- c(set.data$M1[i],set.data$M2[i],set.data$M3[i])
  AvgCounts.data[i,1] <- mean(C,na.rm = T)
  AvgCounts.data[i,2] <- mean(E,na.rm = T)
  AvgCounts.data[i,3] <- mean(Th,na.rm = T)
  AvgCounts.data[i,4] <- mean(M,na.rm = T)
}

```

