---
title: "Image check"
output: html_notebook
---
PURPOSE: This script reads in cell count data from excel spreadsheet

APPROACH: 
1. Read in cell count data from excel spreadsheet
2. Separate samples by tree
3. Go through each sample in each tree
4. Check Yr-1 for each sample: FR yes or no?
5. Enter data in new column

RESOURCES
1. Joining data tables based on a column
https://www.infoworld.com/article/3454356/how-to-merge-data-in-r-using-r-merge-dplyr-or-datatable.html Tutorial video
    data: bit.ly/USFlightDelays
    code: www.infoworld.com/article/3454356
    

I. Read in data from spreadsheet
```{r}
library(readxl)
library(TDPanalysis)
library(lubridate)
library(stringr)
library(dplyr)
library(tidyr)
library(imager)

# clear workspace
rm(list=ls())

fname_current <- "BIG_from 01JULY2020_2.xlsx"
pathname1 <- file.path("~/Documents/Xylogenesis/Project_6 Monson 2018/2. Data/1. Cell Counts/BIG",fname_current)

# [5JAN21] need to insert pathname
data <- read_xlsx(pathname1, 
                  sheet = "data",
                  range = "A1:AF471", 
                  na = "-999")
data <- data.frame(data)
```

```{r}
# "actions" are labelled: "RE-DO","REJECT" or "NA". For "NA" records, there is cell count data
# Parse out samples that need to be Redone or Rejected
data.RR <- subset(data,!is.na(data$actions))

# ~~~~~~~~~~~~~~~~~~  ONLY RECORDS WITH CELL COUNT DATA

# Parse out only samples with cell count data
dataCC.data <- subset(data,is.na(data$actions))
#dataCC.data2 <- dataCC.data[,c(-1,-2,-4)]
#dataCC.data2 <- data.frame(dataCC.data2)
nrec <- length(dataCC.data[,1])

# Convert to data frame
dataCC.data <- data.frame(dataCC.data)

# CONVERT DAYS TO JULIAN DAYS
ndays_month <- c(31,28,31,30,31,30,31,31,30,31,30,31)
ndays_moncum <- cumsum(ndays_month)

doy <- ndays_moncum[dataCC.data$mon_fld-1]+as.numeric(dataCC.data$day_fld)
DOY <- matrix(data=doy,nrow=nrec,ncol=1)
FR_curr <- matrix(data=NA,nrow=nrec,ncol=1)  # Add a column to insert FR presence absence data
FR_tneg1 <- matrix(data=NA,nrow=nrec,ncol=1)  # Add a column to insert FR presence absence data

# add DOY to raw data set
chk1.data <- cbind(DOY,dataCC.data,FR_curr,FR_tneg1)
```


MAIN info for database manipulation from: https://datacarpentry.org/R-ecology-lesson/03-dplyr.html
    SUPP Info 
      1) rowwise info here: https://dplyr.tidyverse.org/articles/rowwise.html
```{r}
# Select specified columns. To select ALL except specified, use "-"
mickey <- select(chk1.data,C1,C2,C3)
notmickey <- select(chk1.data,-C1,-C2,-C3)

# Select rows 
mouse <- filter(chk1.data,mon_fld==7)
mice <- filter(chk1.data,mon_fld==7,day_fld==18) # all samples from July 18

# Mutations and 
# pipes %>% : takes the output of one function and passes it on to the next

# take the mean of the three Yr-1 msmts for each sample (row-wise) and put these in a new column 
chk1.data%>%
  rowwise()%>%
  mutate(YrMinus1=mean(c(rad1,rad2,rad3),na.rm=TRUE))%>%
  head(3)

```

Parse out trees, manually find images and assess
RESOURCES
arrange : https://dplyr.tidyverse.org/reference/arrange.html
```{r}

#ImgCheck <- select(chk1.data,mon_lab,day_lab,yr_lab,Smp_ID,mon_fld,day_fld,yr_fld,IMG_3,IMG_4)

# select only data that was collected in 2019
ImgCheck2019 <- filter(chk1.data,yr_lab==2019)
ImgCheck2019_sort <- arrange(ImgCheck2019,mon_lab,day_lab,IMG_3)

```

Find and open image in RStudio
1. Go to folder
2. Open folder, go to files
```{r}
# List CS folders
#folder <- list.files(path = "~/Documents/Xylogenesis/Project_6 Monson 2018/2. Data/2. Images")

# NEXT: Identify folder of interest by data collection date

# [2APR2021] For now, I will type this in manually
IMGpath <- "~/Documents/Xylogenesis/Project_6 Monson 2018/2. Data/2. Images/Folder_20190601"
ImageList <- list.files(path = IMGpath)
ImageList

image_path <- file.path("~/Documents/Xylogenesis/Project_6 Monson 2018/2. Data/2. Images",folder[1],ImageList[5])

im <- load.image(image_path)
plot(im)
```

